---
title: "Geosoftware II - WiSe 2018/19"
subtitle: "_Enhancing discovery of geospatial datasets in data repositories_"
author: "Daniel Nüst, Prof. Edzer Pebesma"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "wwu-theme.css"]
    nature:
      countIncrementalSlides: false
      beforeInit: ["macros.js"]
      ratio: "16:9"
# use RStudio Addin "Inifinite Moon Reader" for live preview
---

```{r live_preview, include=FALSE, eval=FALSE}
library("xaringan")
xaringan::infinite_moon_reader()
```

```{r pdf_webshot, include=FALSE, eval=FALSE}
library("webshot")
print(knitr::current_input())
system("cat *.Rmd | grep -v '^--$' | grep -v 'ratio:' > print.Rmd")
rmarkdown::render('print.Rmd', 'xaringan::moon_reader')
file_name <- paste0("file://", normalizePath("print.html"))
webshot(file_name, "geosoft2-2018-intro.pdf")
system("rm -r print.Rmd print.html print_files")
```

class: wwu-bluegreen

# Agenda

- Learning objectives
- Teachers
- Project
- Course programme
- Research topics

---
class: wwu-green

## Learning objectives

--

- project management & working in teams

--

- _client & customer_ relationship similar to business projects

--

- handle and extend _large_ open source projects

--

- develop specific features as _small_ independent open source projects

--

- **software development skills for spatial applications**

???

project: GitLab, GitHub, team work

client & customer: (bid, communication, delivery)

FOSS projects: install, maintain, develop further, even in new programming languages

small & large: and gain experiences about the difference

---
class: center,middle,wwu-greenblue

# Teachers

???

## Daniel Nüst

- Dipl.-Geoinf.
- Wiss. Mitarbeiter @ [ifgi](https://www.uni-muenster.de/Geoinformatics/)
- Offene Reproduzierbare Forschung ([o2r](http://o2r.info/))
- Davor: [52°North](https://52north.org/) (consultant,<br />Forschungsprojekte, Softwareentwickler)
- **Free & Open Source Software und Reproducible Research**
- [https://github.com/nuest](https://github.com/nuest)
- [https://orcid.org/0000-0002-0024-5046](https://orcid.org/0000-0002-0024-5046)
- [@nordholmen](https://twitter.com/nordholmen)
- Ultimate Frisbee

---
class: center,middle,wwu-cassis

# Project

--

### or ... "Challenge"

--

## _Enhancing discovery of geospatial datasets in data repositories_

---
class: wwu-bluegreen

# Problem

--

##### more and more research data is be published, incl. geospatial data

--

##### finding the right data and publications get's harder

--

##### recommendation of "similar/related products" commonplace in online shopping, but missing in research data repositories and geospatial catalogues

???

A large value, or potentially a trap for the consumer, in online shopping lies in
recommendation systems.  These systems suggest similar products based on all
or specific customer’s shopping behaviour.  In a similar vein, discovering related
works and similar research is a challenge for scientists.  The challenge applies to
their own work (who is working in my area of expertise or region of interest?)
and finding relevant literature (what publications exist on topic X?).
With the increased requirements by funding agencies to publish data in open
repositories, this project should tap into the wealth of published research data
and provide new ways to connect and discover datasets based on their geospatial
properties.  By extending our catalogue platform and data repository, we except
to gain a significant advantage over competitors.

---
class: wwu-lightgreen

# Solution

--

## &nbsp;

#### _Extend a repository platform or a geospatial catalogue with geospatial similarity measurements for contained records to suggest related records to users._

???

**leverage geospatial aspects of data to connect previously unconnected records

---
class: wwu-lightgreen
# Benefits

--

##### scientists can search better for similar or related work

--

##### preservationists improve data collections

--

##### operators of data repositories or catalogue services can easily integrate similarity measures in their deployments

???

scientists: during research, writing manuscripts, or evaluating papers

---
class: wwu-bluegreen

### Scholarly Computational Research '18

Issue: half time of computational research too small

--

- executable digital notebook ([R Markdown](http://rmarkdown.rstudio.com/), [Jupyter Notebook](https://jupyter.org/)) under version control following best practices of software development

--

- [good file names](http://www2.stat.duke.edu/~rcs46/lectures_2015/01-markdown-git/slides/naming-slides/naming-slides.pdf) and _plain text_ files wherever possible (geo?)

--

- Dokumentation for "future you" as the best collaborateur

--

- **Transparency** & collaboration - "show me", not "trust me"

---
class: center,middle,wwu-lightgreen

# Course programme

Wednesdays, 14:00-16:00 hrs

First meeting @ October 10 2018

---
# Schedule

--

- presentation of **fundamental topics** (individual, groups of two), see [github.com/Geosoft2/geosoft2-2018](https://github.com/Geosoft2/geosoft2-2018)

--

- **invitation to bid** & detailed course schedule

--

- **project work** (3-4 teams): planning, implementation, presentation, peer code review

--

- individual **reports**

???

fundamental topics: help each other understanding the problem and existing domains quicker: handout, presentation
research topics and bid on first meeting (all handouts ready one week in advance)
research topics and questions about bid at second meeting, groups announced

**invitation to bid has a list of functional and non-functional requirements to implement**

project work: groups will be made by us

reports: a couple of pages

---
# Efforts

--

.pull-left[
- 9  ECTS = 270  working  hours
- average workload of ![:inline_highlight](**16.9 hrs per week** per student) (16 weeks)
]

--

.pull-right[
- 24 hours preparation during the semester break
- 16 hours for preparing & attending initial & final presentations
- 12 hours for the individual report and code peer review
- ![:inline_highlight](218 hours) over ~11 weeks for planning & implementation
]

???

**what to expect**

11 weeks implementation does NOT count Christmas holidays

---
# Grading

| What | Work load (hrs) | Grade contribution (%) |
| ---- | ---- | ----|
| initial training, presentations, code peer review & individual report | 52 | 20 |
| planning & prototyping | 32 | 15 | 
| implementation | 186 | 65 |

???

Grading is based on code contributions on the software development platform.
Therefore, we require each student to contribute under their **own account**.

A **fork & pull** development model is highly encouraged because it allows to list the relevant pull requests in a final personal report presenting learning achievements, role in the team, personal contributions in a single PDF document (max. 3 pages).

We strongly encourage vertical splitting of tasks (a student works on UI, backend, and installation) instead of horizontal (one team member does maintenance, another backend, a third frontend) even at the cost of effectiveness.


---
class: wwu-lightgreen

# Research topics

--

- scholarly publishing: history and future [Ilka Pleiser]
- academic search engines [Christin Bösch]
- data publishing requirements in science [Carolin Bronowicz]
- research data lifecycle & best practices [Cornelia Zygar]
- geospatial data formats & libraries\* **[NN]**
- time series data formats & libraries [Katharina Hovestadt]
- FOSS repositories & preservation\* **[NN]**
- publication metadata [Lia Kirsch]
- Geodata catalogues & geospatial metadata [Aysel Tandik]
- metadata extraction [Lukas Jahnich]
- spatial similarity calculation\* [Jannis Fröhlking, Benjamin Dietz]
- agile software development [Daniel Nüst]

???

11 topics

for the * topics please find a partner

Extra topics: Open Science, ..

---
# Next

- register on LernWeb
- quick research into your topic (1 hr) > send Daniel outline of your handout > receive further suggestions
- prepare topic
- submit handout via GitHub before October

--

![:highlight_inline](See you October 10 - **have a good summer!**)
